VFTS682: a confirmed dynamical ejection?

### 2018-04-30 VFTS682.tex
started writing the paper

\renewcommand{\labelitemii}{$\bullet$}
\newcommand{\kms}{{\mathrm{km\ s^{-1}}}}
\newcommand{\Msun}{{\mathrm{M}_\odot}}

@@

How do stars form is one longstanding question in astrophysics
\todo{ref?}. Answering it for massive stars is particularly difficult,
since these are intrinsically rare
\citep[e.g.,][]{salpeter:55,kroupa:01, schneider:18}, evolve fast, and
remain enshrouded in their parent cloud during the whole formation
process. Moreover, observations of young massive stars reveal a
complicated multipliticy structure which requires explanation.

Understanding massive star formation, possibly as a function
of metallicity, is a key question given the present and upcoming
transient survey \citep[e.g., LSST, BlackGem, LIGO/Virgo O3][]{} which will reveal transients associated to their
evolution and death.

@@

The second data release (DR2) from the Gaia satellite
\cite[][]{brown:18,} allows us to test
these hypothesis using one particular star, VFTS682. This star is a
very massive ($\sim150\,M_\odot$) WNh5 star in the 30 Doradus region
of the Large Magellanic Cloud, and it is presently observed at a
projected distance of $\sim$$29$\,pc from the nearest cluster of
massive stars R136 \citep[][]{bestenlehner:11}. Based on the extremely high mass of this star and
its present day apparent isolation, \cite{bestenlehner:11} proposed it
might be a candidate for isolated star formation, if it has not been
ejected from R136 in the past.

In this study, we combine the radial velocity measurements from the
VFTS survey \citep[][]{evans:11} with the proper motion from Gaia DR2
to reconstruct the three-dimensional velocity of VFTS682, and test the
null hypothesis that this star was ejected from R136. \todo{check the
  following} Our results indicate that R136 is the likely origin of
this star, and therefore isolated star formation is \emph{not}
required to explain it. However, we find a discrepancy between the
apparent age of VFTS682, it's kinematic age, and the age estimates for
R136.

In \Secref{sec:sample} we describe how we select stars in R136 and in
the surroundings of VFTS682 itself to remove the mean motion of the
region. \Secref{sec:results} presents our main findings. We conclude by
discussing the implications for theories of
star formation, N-body interactions, and binary evolution in
\Secref{sec:discussion}.

\section{Gaia DR2 data selection}
\label{sec:sample}
VFTS682 is labeled in the Gaia DR2
catalog\footnote{\url{https://gea.esac.esa.int/archive/}} with the
source id 4657685637907503744. For this star, we retrieve from the
Gaia catalog its position in right ascension (RA) and declination (DE)
in the IGCS frame\cite[][]{brown:18}\todo{some 2016 Gaia paper?}, its
proper motion components ($\mu_\mathrm{RA}$, and $\mu_\mathrm{DE}$,
respectively). For the radial velocity of VFTS16 and of the 30 Doradus
region as a whole, we instead use the VFTS data
as quoted in \cite{bestenlehner:11}. Our main conclusions do not
depend on the detailed value of the radial velocity.\todo{improve
  phrasing} The values of all these quantities that we use are listed
in \Tabref{tab:vfts682}.

\begin{table}[tbp]
  \centering
  \begin{tabular}[htbp]{l|c|c}
    Parameter & Value & Source\\ \hline\hline
    RA \hfill[degree] &  84.73 $\pm$  0.036 & \multirow{4}{*}{Gaia DR2}\\
    DE \hfill [degree] & -69.07 $\pm$  0.05  & \\
    $\mu_\mathrm{RA}$  \hfill[$\mathrm{mas\ yr^{-1}}$] & 1.84 $\pm$ 0.07 & \\
    $\mu_\mathrm{DE}$  \hfill[$\mathrm{mas\ yr^{-1}}$] & 0.78 $\pm$ 0.08& \\
    $\delta v_\mathrm{rad}$  \hfill[$\kms$] & -45 $\pm$ 25 & \cite{bestenlehner:11}\\
    \hline
  \end{tabular}
  \caption{Astrometric parameters for VFTS682}
  \label{tab:vfts682}
\end{table}

@@

\section{The kinematics of VFTS682}

\begin{figure*}[htbp]
  \centering
  \includegraphics[width=\textwidth]{./figures/main_plot}  
  \caption{position and projected relative velocity to R136}
  \label{fig:main}
\end{figure*}

>>>>

\renewcommand{\labelitemii}{$\bullet$}
\newcommand{\kms}{{\,\mathrm{km\ s^{-1}}}}
\newcommand{\Msun}{{\mathrm{M}_\odot}}

@@

How do stars form is one longstanding question in astrophysics \cite{lada:03}
\todo{ref?}. Answering it for massive stars is particularly difficult,
since these are intrinsically rare
\citep[e.g.,][]{salpeter:55,kroupa:01, schneider:18}, evolve fast, and
remain enshrouded in their parent cloud during the whole formation
process \todo{ref}. Moreover, observations of young massive stars reveal a
complicated multipliticy structure which requires
explanation \citep[][]{kobulnicky:07,mason:09,sana:11,sana:12,kiminki:12,chini:12,kobulnicky:14,almeida:17,demarco:17}. Understanding massive star formation, possibly as a
function of metallicity, is a key question given the present and upcoming
transient survey \citep[e.g., LSST, BlackGem, LIGO/Virgo O3][]{} which
will reveal transients associated to massive stars
evolution and death.

@@

The second data release (DR2) from the Gaia satellite
\cite[][]{brown:18,} allows us to test
these hypothesis using one particular star, VFTS682. This star is a
very massive ($\sim$$150\,M_\odot$) WNh5 star in the 30 Doradus region
of the Large Magellanic Cloud, and it is presently observed at a
projected distance of $\sim$$29$\,pc from the nearest cluster of
massive stars R136 \citep[][]{bestenlehner:11}. Based on the extremely high mass of this star and
its present day apparent isolation, \cite{bestenlehner:11} proposed it
might be a candidate for isolated star formation, or a runaway ejected
from R136 in the past.

Massive stars can in principle be ejected from R136 as a consequence of dynamical
interactions \citep[][]{poveda:67,leonard:91, evans:10, fujii:11,
  allison:12, oh:16}, or by the disruption of a binary by the first
core-collapse supernova
\citep[][]{zwicky:57,blaauw:61,dedonder:97,eldridge:11,renzo:18}. However,
R136 has an estimated age of $\lessim2$\,Myr \todo{ref}, which is
shorter than the shortest stellar lifetime of about $\sim$3\,Myr
\citep[e.g.,][]{zapartas:17}, so one would not expect the binary
disruption scenario to be relevant for this cluster.

In this study, we combine the radial velocity measurements from the
VFTS survey \citep[][]{evans:11} with the proper motion from Gaia DR2
to reconstruct the three-dimensional velocity of VFTS682, and test the
null hypothesis that this star was ejected from R136. \todo{check the
  following} Our results indicate that R136 is the likely origin of
this star, and therefore isolated star formation is \emph{not}
required to explain it. However, we find a mild discrepancy between the
apparent age of VFTS682, its kinematic age, and the age estimates for
R136.

In \Secref{sec:sample}, we describe how we select stars in R136 and in
the surroundings of VFTS682 itself to remove the mean motion of the
region. \Secref{sec:results} presents our main findings. We conclude by
discussing the implications for theories of
star formation, N-body interactions, and binary evolution in
\Secref{sec:discussion}.

\section{Gaia DR2 data selection}
\label{sec:sample}
VFTS682 is labeled in the Gaia DR2
catalog\footnote{\url{https://gea.esac.esa.int/archive/}} with the
source id 4657685637907503744. For this star, we retrieve from the
Gaia catalog its position in right ascension (RA) and declination (DE)
in the IGCS frame\cite[][]{brown:18}\todo{some 2016 Gaia paper?}, its
proper motion components ($\mu_\mathrm{RA}$, and $\mu_\mathrm{DE}$,
respectively). For the radial velocity of VFTS682 and of the 30 Doradus
region as a whole, we instead use the VFTS data
as quoted in \cite{bestenlehner:11}. Our main conclusions do not
depend on the detailed value of the radial velocity.\todo{improve
  phrasing} \Tabref{tab:vfts682} lists the values adopted throughout
this work for each of these quantities.

\begin{table}[tbp]
  \centering
    \caption{Astrometric parameters for VFTS682. The peculiar radial
    velocity $\delta v_\mathrm{rad}$ is obtained as the difference
    between the average radial velocity of the 30 Doradus region
    ($270\pm10\kms$) minus the radial velocity measured from the HeII $\lambda4686$
    line for VFTS682 ($315\pm15\kms$).}

  \begin{tabular}[htbp]{l|c|c}
    Parameter & Value & Source\\ \hline\hline
    RA \hfill[degree] &  84.73 $\pm$  0.036 & \multirow{4}{*}{Gaia DR2}\\
    DE \hfill [degree] & -69.07 $\pm$  0.05  & \\
    $\mu_\mathrm{RA}$  \hfill[$\mathrm{mas\ yr^{-1}}$] & 1.84 $\pm$ 0.07 & \\
    $\mu_\mathrm{DE}$  \hfill[$\mathrm{mas\ yr^{-1}}$] & 0.78 $\pm$ 0.08& \\
    $\delta v_\mathrm{rad}$  \hfill[$\kms$] & -45 $\pm$ 25 & \cite{bestenlehner:11}\\
    \hline
  \end{tabular}
  \label{tab:vfts682}
\end{table}

@@

The 
\section{The kinematics of VFTS682}
\label{sec:results}

\begin{figure*}[htbp]
  \centering
  \includegraphics[width=\textwidth]{./figures/main_plot}  
  \caption{position and projected relative velocity to
    R136. \todo{load pretty picture on background, check scale, clean
      foreground stars with parallax, add cone of uncertainty}}
  \label{fig:main}
\end{figure*}

\section{Discussion}
\label{sec:discussion}
